\appendix
\section{A Little History of Discovering Moving Objects}
\label{history}

{\it Contributed by Jeff Larsen}

To provide a bit of background, the task of discovering unknown asteroids has a very
lengthy history, involving human, electronic, and computerized actors.

From the beginning the primary detectable feature of an asteroid is
it's motion across the sky as time progresses -- which makes the
underlying orbital parameters of critical relevance to the continued
tracking of the object.  The first asteroid, Ceres, was visually
detected by Piazzi \citep{1802QB378} through its motions
against the background star as he attempted to verify the position of
a star in a published catalog.  Without knowing how orbits were
goverened, Piazzi's initial tracking effort involved constant repeated
night-to-night observations which he was only able to maintain for a
limited time due to his health and the motion of the asteroid into the
daytime sky.  It fell to Carl Friedrich Gauss to develop his method of
orbit determination before Ceres could be recovered again thus firmly
establishing the mathematics of orbits as a required prerequisite.

Optical observations had their limitations, however, and the archival
record created by the application of photography to observations was
of critical importance to scientific demands of repeatability.  The
first photographic discovery of an asteroid, (323) Brucia, was by Max
Wolf in 1891 \citep{1892AN} as he recognized the
non-sidereal rate trails on his photographs for the asteroids they
were.  Wolf became a prolific discoverer of asteroids (almost 250
total) through photography and as such developed the standard
detection technique used for the next century.

Photographic sensitivity is determined to a large part through
exposure time and asteroid motion to first order is a function of its
distance so two main types of photographic survey methodology were
used.  Surveys looking for closeby and fast near-Earth asteroids (like
those of \cite{1988NASTM4041...52S}) used non-sidereal trailed motions
and generally shorter exposure times.  A discovered trail of any
length in the and image developed in an observatory darkroom would
lead to additional exposures of the asteroid being quickly acquired
and the data required to secure the orbit.  The second technique,
useful for very slow moving asteroids that did not appear as a trailed
images was to take several exposures and then manually alternate
between them rapidly, thus depending on the well-developed human
motion sense to discover the moving asteroids.  Machines were built
for this purpose, such as the blink comparator.  Besides its use for
stellar proper motions, this technique's most well known success was
in the discovery of Pluto by Tombaugh \citep{1960S&T....19..264T}.
The blink comparator was useful for scientific studies of asteroids in
the main asteroid belt such as the seminal Palomar-Leiden survey
\citep{1970A&AS....2..339V}.  Keeping the exposure times short was
important to keep the asteroid images starlike so their isophotal
diameter could be used inaccurate brightness calibration.  This
allowed for the rapid discovery of asteroids for the time: 2400
discoveries in 11 nights of observation.

The advent of the CCD was applied to asteroids early, with the long
readout times (and consequential 50\% duty cycle) minimized through
the process of drift scannings.  Gehrels and the Spacewatch Project
\citep{1990ASPC....8...51G} pioneered the use of CCD's in the
detection of asteroids, making the first CCD detection of the
near-Earth asteroid 1989 UP and the first discovery as well (1990 SS).
Spacewatch went on to built an advanced real time asteroid motion
detection program, MODP \citep{1991AJ....101.1518R}.  Drift scanning
was made unnecessary by the advent of large format cameras used in a
step-and-stare mode such as that of the NEAT project
\citep{1999AJ....117.1616P}.  With the introduction of custom
frame-transfer CCD's and high-capacity computer processing, the LINEAR
project \citep{2000Icar..148...21S} and the Catalina Sky Survey
\citep{2007IAUS..236..323L} were able to progress handily towards the
Spaceguard goal.

As with photography, there are multiple techniques which can be
applied to CCD images to discover their asteroids.  Most asteroid
detection software uses a "moving target indicator" approach in that
CCD images are searched automatically for their objects who are then
compiled as a detection list.  By filtering out objects which did not
move and searching for asteroid-like motions in the unmatched lists,
motion candidates are created.  However, a more exotic approach called
Matched Filter Processing \citep{2005AJ....130.1951G} can be applied
that has several advantages.  Images are coadded rapidly along
hypothesized motion vectors.  If the flux of an object appears to grow
after coaddition, it becomes a candidate moving object with the motion
vector already determined.  Any technique to search for moving objects
below the single-image limiting magnitude becomes computationally
expensive with a large number of possible motion vectors, but has the
advantage of being able to detect fainter objects in the same set of
images compared to the moving target indicator approach.  Variants on
these techniques have been
used successfully in searches for distant objects such as those in the
Kuiper Belt \citep{BG, MB} and has been used to do NEA searches over
large portions of the sky where the asteroid is uncertain but has a
relatively small range of possible velocities
\citep{2005AJ....130.1951G}.

The advent of large-format all sky surveys have radically changed the
detection requirements both in terms of techniques for imagery and
software, eventually leading to the PS and LSST MOPS software. 



\section{About the KD-Tree Library}
\label{kdTreeLib}

The findTracklets, collapseTracklets, and linkTracklets algorithms
used in DayMOPS all require the use of KD-Trees for different types of
data.  FindTracklets needs an (RA, Dec) tree of detections,
collapseTracklets needs an (RA, Dec, velocity, angle) tree of
tracklets, and linkTracklets needs a specialized (RA, Dec, RA vel,
Dec vel) tree of tracklets.  Other ``helper'' tools not described in
this document have been constructed using trees over various types of
data (such as collections of orbits).

To suit the needs of our various algorithms, and to prepare for the
possibility of new algorithms, we have created our own KD-Tree
library.  The design is intended to fulfill the following needs:

\begin{itemize}
\item Hold spatial points in arbitrary-dimensional space, with each point mapped to an arbitrary piece of non-spatial data
\item Deal with spatial axes which hold real-number values
\item Deal with spatial axes which hold degree or radian values on a circle
\item Deal with pairs of axes which may describe points on a sphere
\item Allow range searching in circular, spherical or flat Euclidean space
\end{itemize}

The {\tt KDTree} class we have created suits all of these needs.
These features are sufficient for almost all our algorithms.  However,
linkTracklets uses trees in unusual ways and its performance is
closely related to the characteristics of the tree it uses; in order
to allow linkTracklets' special needs, we also created a class called
{\tt TrackletTree} used only for linkTracklets.  Both {\tt KDTree} and
{\tt TrackletTree} derive from a common base class {\tt BaseKDTree}
which implements memory management and other common features.

\subsection{Representing Data Items: PointAndValue Class}

Most DayMOPS algorithms deal with identifying groups of data items
based on their spatial location in some coordinate system.  To
represent a data item and its spatial location, we have created a
template class called {\tt PointAndValue}.  A {\tt PointAndValue}
holds a point in an arbitrary-dimensional coordinate system,
represented as a {\tt std::vector} of {\tt double} values.  The
coordinate system may be arbitrary-dimensional, and may also be
heterogenous; that is, some axes may be Euclidean (having arbitrary
values) but others may be circular, or two axes may describe a sphere.
The ``value'' may be anything - this is a templatized data type.  In
our code, it usually holds an integer, representing the location of a
detection or tracklet in an array.  Using a {\tt PointAndValue}
instance, one can represent any data object (the ``value'' of
arbitrary type) and a location of that data object.

\subsection{Tree Construction and Searching}

{\tt KDTree} instances are constructed from a {\tt std::vector} of
{\tt PointAndValue} instances.  The $k$ value (the dimensionality of
the tree) must be specified at construction-time.  Each ``point''
vector used to build the tree must have at least $k$
elements, though the software will allow ``points'' with $>k$
elements, in which case only the first $k$ are used.  

When providing the data for points used in tree construction, it is up
to the user to ensure that any value intended to represent a
measurement in degrees must fall along $[0,360)$.  Failure to do so
will lead to an exception later.

If you wish to treat your spatial points as all Euclidean, it is
possible to perform a conventional range search, which searches for
all points in the tree within a given distance of a query point using
{\tt KDTree::rangeSearch}.  However, if your axes use differing units,
or one or more describe points on a circle, range search will not be
suitable, as it does not handle wrap-around in degree units.  This
type of searching is provided by the library, but is not used within
DayMOPS.

In order to deal with data points which may not be in a Euclidean
space, the library supports rectangular (or hyper-rectangular)
searches on mixtures of Euclidean and/or
circular axes.  A query point is provided in each axis, and a query
range around that point; any point which is within the query range
of the query point in each axis is returned.  This is suitable when
searching, for example, for motion vectors, which have a velocity and
angle of motion; specify a range of velocities, and a range of angles,
and the tree library will find motion vectors which fall within the
given range.  Wraparound will be handled intelligently (that is, the
code will recognize that .1 and 359.9 degrees are separated by .2
degrees, not 359.7.)  To use this type of searching, use the {\tt
KDTree::hyperRectangleSearch} method. Note that it will be necessary
to inform the tree of which axes are Euclidean, and which are
circular.  Currently, the only supported units for circular
coordinates are degrees, and degree measures must fall along $[0,
360)$.  This is the type of searching used in the {\tt
collapseTracklets} implementation, which searches for similar points
in (RA, Dec, velocity and motion angle).  Unfortunately, this type of
searching considers each axis independently, which can be problematic
when dealing with a pair of axes which describe points on a sphere,
where polar distortions may occur.  

To deal with coordinates on a sphere and intelligently account for
both wraparound and polar distortion, the library provides the
function {\tt KDTree::RADecRangeSearch}.  This function performs a
range search on the surface of a sphere; it also allows rectangular
searching of other axes if needed.  The types of the axes should be
specified at searching time as in {\tt KDTree::hyperRectangleSearch},
if any.  This is the search function used by findTracklets, which
represents points as simple RA, Dec coordinates.  It would probably be
wise to change collapseTracklets to use this function as well, though
this has not yet been done.


\subsection{TrackletTree}

The linkTracklets algorithm uses KD-Trees in an unusual fashion, and
is highly sensitive to their construction.  It also requires the
spatial regions held by each tree node to be extended to account for
error bars on the tracklets.  For these reasons, linkTracklets does
not use the normal {\tt KDTree} class used by other algorithms, but a
sibling class called {\tt TrackletTree}.  These classes share the
common ancestor {\tt BaseKDTree}, which handles memory management, but
otherwise they share relatively little.  {\tt TrackletTree} calculates
the extents of the nodes differently than {\tt KDTree}, and it
implements no range searching.  

{\tt KDTree} and {\tt TrackletTree} hold references to the head of the
actual tree nodes, implemented by {\tt KDTreeNode} and {\tt
TrackletTreeNode}.  In {\tt KDTree} these child nodes are not visible
to the user; all communication is done with the {\tt KDTree}, which in
turn communicates with the {\tt KDTreeNode}s.  This allows easier
memory management, as an outside tool cannot hold a reference to a
dynamically-allocated child.  However, in linkTracklets these children
must be visible to linkTracklets, as it explicitly traverses the tree.  

{\tt TrackletTreeNode} also has a few methods for studying performance: \\
{\tt TrackletTreeNode::addVisit} and {\tt TrackletTreeNode::getNumVisits} can
be used to count the number of times a node is examined by
linkTracklets.  Eventually, it may be wise to remove these in order to
conserve memory.


\subsection{BaseKDTree and Memory Management}

{\tt KDTree} and {\tt TrackletTree} represent a whole tree, not the
individual nodes which make up the tree.  {\tt KDTree} then provides
an interface to the querying of the tree, while {\tt TrackletTree}
provides accessors allowing an outside class to see the actual tree
nodes.  

Both classes derive from {\tt BaseKDTree} which handles the mundane
work of a C++ class: construction, copying, and detruction, etc. {\tt
BaseKDTree} handles copying of a tree by incrementing a refcount on
the child nodes; once those nodes hit a refcount of zero, then the
nodes are deleted.  This approach works well, provided no outside
class holds on to references to a tree's nodes after destroying the
tree itself.




\subsection{Needed Improvements}
The KD-Tree nodes at present use quite a bit of memory because they
use {\tt std::vector} where a set of pointers or C-style array would
probably be better-suited.  This is particularly noteworthy in the
case of {\tt TrackletTreeNode}, since trees given to linkTracklets can
be very large.



